---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
setwd("~/sandbox/repdata-008/RepData_PeerAssessment1")
if(!file.exists("activity.csv") {
  stop("Error: Could not find data file. Please check if activity.csv exists.")
}

data <- read.csv("activity.csv")

```
## What is mean total number of steps taken per day?

```{r}
# Libraries
library(plyr)
library(ggplot2)
library(scales)   # For pretty_breaks()

# Print mean_steps to file.
png(filename = "./mean_steps.png")

sumdata <- ddply(data, .(date), summarise, sum_steps = sum(steps, na.rm=TRUE))
meansum <- mean(sumdata$sum_steps)
mediansum <- median(sumdata$sum_steps)

# Setup the plot
p1 <- ggplot(data=sumdata, aes(x=sum_steps)) + geom_histogram(binwidth=2000,colour="black",fill="white") + geom_vline(aes(xintercept=mean(sum_steps, na.rm=TRUE)), color="red", linetype="dashed", size=1) + geom_vline(aes(xintercept=median(sum_steps, na.rm=TRUE)), color="blue", linetype="dashed", size=1)

# Make it pretty
p1 + ylab("Count") + xlab("Total Steps") + geom_text(aes(meansum - 4000,12,label=paste("Mean ~ ",as.integer(meansum)),color="red")) + geom_text(aes(mediansum + 6000, 12, label=paste("Median = ",mediansum)),color="blue") + theme(legend.position = "none")

# Close device
dev.off()
```

The mean total number of steps taken per day are `r meansum`.



## What is the average daily activity pattern?

A time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r}
# Calculate the 5 minute interval average number of steps
fiveInterval <- ddply(data, .(interval), summarise, avgSteps = mean(steps, na.rm=TRUE))

# Print avg_interval to file.
png(filename = "./avg_interval.png")

p2 <- ggplot(data=fiveInterval, aes(x=interval)) + geom_line(aes(y=avgSteps))
p2 +  ylab("Average Steps for all days") + xlab("5-minute Interval")

# Close device
dev.off()

```


```{r}
# Determine the max value interval and save it
maxInt_idx <- which.max(fiveInterval$avgSteps)
```

The 5-minute interval `r fiveInterval[maxInt_idx,1]`, on average across all the days in the dataset, contains the maximum number of `r fiveInterval[maxInt_idx,2]` steps.


## Imputing missing values

```{r}
# Using the previous calculated fiveInterval averages, fill in missing data

data_more <- as.data.frame(data)
data_more[is.na(steps)] <- 
  
  data_more[with(data_more,is.na(steps)),]$steps <-

  with(data_more, ifelse(is.na(steps)), match(fiveInterval$interval, 
                                              
  # Lookup table for missing value
  fiveInterval[which(fiveInterval$interval == data_more$interval),2]                                            
```


## Are there differences in activity patterns between weekdays and weekends?
